Uppangala tree functionnal: RLQ draft

### 2016-07-08 inst/doc/RLQ.Rmd
v0.1.9017 RLQ.Rmd + plan

---
title: 'Uppangala tree functionnal traits: RLQ and Fourth cornr analysis'
author: "Sylvain Schmitt"
date: '`r Sys.Date()`'
output: pdf_document
bibliography: refs.bib
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
```

# Traits

## Measured traits

Table 1: Functional traits measured. *Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree development.*

@@

Uppangala measured functional traits vary in the ranges of tropical species values that we can find in literature. We can notice in table 1 that wood density has a relatively high mean as we could expect in an old untouched forest. Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010], and can be as well explained by environment condition. For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011]. We could also have added a sixth trait, called leaf tissue density (LTD) with the formula: $LTD=Leaf dry weight/Thickness*Leaf area$. LTD is highly correlated to LDMC [@hodgson_is_2011], but can still bring extra information **To be discussed**.

## Traits distribution and variation

@@

Some outliers are still to be corrected in individual leaf trait value, but thickness outliers and extreme values are due to too close secondary veins. Table 2 shows a intra-individual variation lower than intra-specific and inter-specific variation, indicating that leaves are even in morphology inside an individual but can start to show some variation between trees of the same species. More specifically, whereas traits like thickness and wood density seems to stay stable in intra-individual and intra-specific, leaf area shows a high coefficient of variation for both intra-individual and intra-specific values.

## Covariations in traits

```{r 3 PCA PFT}
library(vegan)
pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_ind, cor = T)
lambda <- pca$sdev * sqrt(nrow(pca$scores))
pca.var = apply(pca$score, 2, var)
PCAscores <- as.data.frame(pca$scores)[1:2] # Adding PCA axis in PFT values
names(PCAscores) <- c('WES', 'LES')
PCAscores$ID <- row.names(PCAscores)
plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
     xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
# as.factor(Species[PFT_ind[row.names(pca$scores),2],6])
par(new=T)
Rot <- t(t(pca$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
       text.col = c("green", "firebrick"))

arrows (rep(0,nrow(pca$loadings)),rep(0,nrow(pca$loadings)),Rot[,1],Rot[,2],lwd=2)
text (Rot[,1:2], pos = c(4, 2, 4, 2, 3), rownames(Rot),cex=1.2)
Hmisc::subplot(function(){
  b <- barplot((pca.var / sum(pca.var)), las = 2, ylim = c(0,range(bstick(7))[2]),
          cex.names = 0.7, axes = T, names.arg = NA) ;
  lines(b, bstick(length(pca.var)), col = 'red') ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.75,1), from='npc'),
  y=grconvertY(c(0.05,0.30), from='npc')
)
axis (3)
axis (4)
PFT_ind$ID <- row.names(PFT_ind)
PFT_ind <- merge(PFT_ind, PCAscores, all = T)
row.names(PFT_ind) <- PFT_ind$ID
PFT_ind <- PFT_ind[-which(names(PFT_ind) == 'ID')]
```

@@

We realized a principal component analysis on individual trait means (see figure 1), where first axis (PCA1) covary mainly with leaf thickness, dry matter content and wood density. Whereas second axis (PCA2) covary mainly with specific leaf area representative of the leaf economic spectrum [LES, @wright_worldwide_2004]. PCA1 might be associated with the wood density covariation to the wood economic spectrum [WES, @chave_towards_2009]. Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us). Our results integrate less traits (5/14), and less species (92/758), but still capture the same strategy axes.

We can see that as wood density and specific leaf area are orthogonal following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$). Leaf area is almost opposed to wood density and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007. This relationship could be explained by an equilibrium in tree hydraulics (Coomes et al, 2008). Finally, we found the positive relation suggested by @diaz_global_2015 between leaf dry matter content and wood density (Spearman's $rho = 0.24$, $p-value < 0.001$).

# Environmental filters

## Environment and trees

We collected trait information on 92 over 117 (79%) species focusing in priority on abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates (*Warning : one of our 92 species is not referenced yet among currently known Uppangala species, Agrosystachis indica, one of 117 species is not removed yet, Toona ciliata.*). Thus we generated data for 98.8% of the tree individuals present in Uppangala plot.

```{r 4 env-com}
library(raster)
Env <- genEnv()
Env$Rivers <- readAll(Env$Rivers)
Trees <- genTrees()
Species <- genSpecies()
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
Trees$SpCode <- tolower(Trees$SpCode)
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
XY <- XY[-which(is.na(XY$X)),] # Removing NA
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
com@data$id <- paste0('C', com@data$id)
com@data$DEM <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'DEM'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'aspect'))
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$SW <- sin((com@data$Aspect + 225) / 180 * pi)
com@data$NE <- cos((com@data$Aspect + 225) / 180 * pi)
com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$topo <- as.factor(with(com@data, topo_classes(Slope, Curvature, Aspect)))
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
# com@data$Height <- aggregate(Trees$height, list(Trees$com), mean, na.rm = T)[,2]
rm(Deriv, XY)
```

To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river. Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$. Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect in degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008]. Distance to the nearest potential river was caluclated under ArcGis (cite).

In order to avoid multi-colinearity in future regressions, we reduced our explanatory variables via a principal component analysis, see figure 2. Curvature, profile curvature, plan curvature, and distance to potential rivers covary together and with the principal component analysis first axis (PCA1). Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables. As a conclusion, we choose to reduce our variable to 4: curvature to represent the water flow paths and soil stabilization areas, wetness to highlight water accumulation areas, southwesterness (SW) to represent monsoon winds [@bunyan_effect_2015] and slope.

>>>>

---
title: 'Uppangala tree functionnal: RLQ draft'
author: Sylvain Schmitt sylvain.schmitt@agroparistech.fr
output: pdf_document
bibliography: refs.bib
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
library(pander)
opts_chunk$set(
  echo = F, message = F, warning = F, fig.height = 8, fig.width = 8,
  cache = T, cache.lazy = F)
```

# Abstract

Keywords: functionnal trait, tropical evergreen forest, Western Ghats, India...

# Introduction

# Material and methods

## Study site

The study was carried out in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), in the 10-ha "Uppangala plot" named after the nearest village. The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996. The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergreen forest type [@pascal_wet_1988]. Climate and geology of the 10-ha plot can be considered as homogeneous. General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the main local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998]. Each tree girth is monitored by a two-year census.

## Sampling

We collected trait information on 92 over 117 (79%) species focusing in priority on most abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates. Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals. Thus we generated data for 98.8% of the tree individuals present in the Uppangala plot. For each individual we collected 6 mature shadow dry leaves and one branch sample with a diameter close to 1cm following recommendation from @perez-harguindeguy_new_2013. Individuals and leaves were selected regarding their light condition, development stage and sanitary state for consistancy. Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling. Leaf and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours. For some individuals we were unable to collect a correct branch sample (58 over 267 individuals representing 12 species over 91).

Leaf petioles, rachis and petiolules were removed for all measurements. Leaf fresh weights, area and thickness were measured, using a scanner [cite] and a microgauge [cite]. Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying process was stopped after dry leaf weights were stable in time. Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias. Branch samples bark and pit were removed in order to keep only branch wood with no soft tissues. Branch wood samples were sink in water to evaluate their volume by water displacement. Samples were then dried for 48 hours at 80 degrees celsius and weighted.

## Traits

With leaf and branch wood samples we calculated five functionnal traits: leaf thickness (Thick) measured in micrometer, leaf area (LA) measured from scans in $mm^2$, leaf dry matter content (LDMC) as the leaf dry weight in $mg$ divided by the leaf fresh weight in $g$,  specific leaf area (SLA) as the leaf area in $m^2$ divided by the leaf dry mass in $kg$, and branch wood density (WD) as the wood sample volume in $cm^3$ divided by the wood sample dry weight in $g$ (see table 1). Functional traits can be linked to trade-off in tree development strategy [@baraloto_decoupled_2010, and see table 1], and can be as well explained by environment condition. For instance, SLA is a soil fertility - shade index whereas LDMC will only be linked to soil fertility [@hodgson_is_2011].

Table 1: Functional traits measured. *Traits with their abbreviation, range of values in the sampling, mean and standard deviation, standard unit, and associated trade-off in tree development from @baraloto_decoupled_2010.*

@@

## Environmental variables

To study environment effect on our traits we choose 8 topographical variables representing abiotic factors: digital elevation model, slope, curvature, plan curvature, profile curvature, wetness, southwesterness (SW, aspect in direction of south-west), and distance to the nearest potential river. Digital elevation model is issued from a LIDAR campaign on Uppangala plots with a resolution of $1*1 m^2$. Slope, curvature, plan curvature, profile curvature, wetness and southwesterness ($cos(aspect~in~degres + 225)$) are all derived from the digital elevation model with the RSAGA package [@brenning_statistical_2008]. Distance to the nearest potential river was caluclated under ArcGis (cite).

To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density. Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$. Both basal area and stem density are calculated with 2013 girth inventory from Uppangala plot on stem with a diameter above 10 centimeters at breast height ($DBH>10$). Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$). So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).

## Analyses

### Covariation in traits

Traits variation was first investigated in three levels with the coefficient of variation ($cv=sd/mean$) between leaves of the same individual (intra-individual), between trees of the same species (intra-specific), and between species of the Uppangala plot (inter-specific). For the rest of the statistical analysis, we calculated the mean value of each 5 traits for our 92 species. To test the correlation and covariation among traits we performed a principal component analysis (PCA). We investigated peculiar traits covariation highlighted with the PCA with a Spearman's rank correlation test.

### Environment definition

In order to avoid multi-colinearity in future analysis, we reduced our abiotic factors explanatory variables via a principal component analysis (PCA), by keeping less correlated variables with the most ecological meaning to our judgement. In order to see if or if environmental biotic factors were the results of functionnal traits we used structural equation modelling. Three latent variables were defined: *Abiotic environment* defined with selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits. Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*). Models were compared with a Chi Square Difference test on Aikaike's Criterion Indice (AIC).

### Community analysis

Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats. For all communities, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes. We used the following formula to compute each community weighted parameter (CWP):
$$CWP = \sum_{k=1}^{nj} W_{k,j} * P_{k,j} ~ (1)$$ 
where $nj$ is the number of species sampled in community $j$, $W_{k,j}$ is the weight of species $k$ in community $j$ used to compute CWP (relative abundance, presence-absence, or relative basal area), and $P_{k,j}$ is the parameter of species $k$ in community $j$ (mean or the variance). In addition 5 functionnal diversity indices (FD) weighted by abundances were computed: functional richness (FRic), functional evenness (FEve), functional divergence (FDiv) and Rao's quadratic entropy (Q) (@villeger_new_2008, @botta-dukat_raos_2005). Spatial variograms were used to investigate our CWM and CWV spatial organization. CWM, CWV, and FD spatial autocorrelation (SAC) was also assessed by Moran's I statistic.

Links between both CWM and CWV with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables. Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables. A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix). Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance. Same procedure was also applied to functionnal diversity indices (FD) from our communities. Finally spatial autocorrelation (SAC) of residuals was investigated to assess selected models.

### Fourth-corner analysis

In addition of our community analysis by regression, we conducted both RLQ and fourth-corner analyses to test the effect of environments on Uppangala tree functionnal traits following @dray_testing_2008. Both matrices R (`Environment*Sites`) with selected features and Q (`Traits*Species`) were reduced by a principal component analysis. We used a Monte-Carlo test to compare fourth-corner and RLQ results to 999 nul models built with the model *m6* proposed by @ter_braak_improved_2012. Fourth corner analysis was tested with 3 types of L matrices ($Sites*Species$) with species relative abundance, presence absence or relative basal area in each sites; whereas RLQ analysis was only conducted with relative abundance.

### Phylogenetic signal

Phylogenetic tree of our 92 species was built with `S.PhyloMaker` function implemented by @qian_updated_2016-1. Phylogenetic autocorrelation was investigated with Morans'I indice on 60 species that had information for all 5 functionnal traits, the two PCA axes and in the phylogenetic tree. Phylogenetic signal structure was assessed by a tree tip distances correlogram. In addition, $\beta st$ and $\pi st$ metric proposed by @hardy_characterizing_2007 were used to investigate phylogenetic signal spatial organization with the null model *"1s"*" suggested by @hardy_testing_2008 to see if we were under phylogenetic clustering or overdispersion. Finally phylogenetic association to habitat was also studied by an RLQ analysis with tree tip distances as the Q matrix reduced by a principal coordinate analysis. RLQ result with phylogenetic distance matrix was compared to 999 nul models were phylogenetic tree tips were randomized.

All analysis were conducted in the R 3.3.0  free software environment for statistical computing and graphics [@r_core_team_r:_2016] using the pacakge 
`ade4` [@dray_ade4_2007], 
`lavaan` [@rosseel_lavaan:_2012], 
`FD` [@laliberte_distance-based_2010], 
`gstat` [@pebesma_multivariable_2004], 
`ape` [@paradis_ape:_2004], 
`phylosignal` [@keck_phylosignal:_2015] 
and `spacodiR` [@eastman_spacodir:_2013].

# Results

## Are functionnal composition of communities linked to the environment ?

@@

Trait coefficients of variation analysis reveals a general increase of functionnal trait variation among scales (from intra-individual to inter-specific, see table 2). Whereas LA shows already high variablilty at intra-individual level almost as strong as the intra-specific variation ($CV_{individual} = 0.20$ and $CV_{species} = 0.24$).

The multivariate analysis of trait correlation shows WD associated with structural leaf traits orthogonal to the SLA (see figure 1). The first PCA axis (PCA1) covary mainly with Thick, LDMC and WD; opposing thin and big leaves to dense leaves and wood. Whereas the second PCA axis (PCA2) covary mainly with SLA representative of the leaf economic spectrum [LES, @wright_worldwide_2004]. PCA first axis might be associated with the WD covariation to the wood economic spectrum [WES, @chave_towards_2009]. Results are similar to the one found by @fortunel_leaf_2012 where PCA1 and PCA2 were respectively interpreted as leaf and wood economic spectrum (inverted compare to us). Our results integrate less traits (5 against 14), and less species (92 against 758), but still capture the same strategy axes.

Figure 1 shows WD orthogonal to SLA following the results of @baraloto_decoupled_2010 suggesting that the two economic spectrum were decoupled (Spearman's $rho = -0.04$, $p-value > 0.1$). LA is almost opposed to WD and is effectively negatively correlated to it (Spearman's $rho = -0.033$, $p-value < 0.001$) as found by @wright_relationships_2007. Finally, we found the positive relation suggested by @diaz_global_2015 between LDMC and WD (Spearman's $rho = 0.24$, $p-value < 0.001$).

```{r 3 RLQ}
library(vegan)
library(ade4)
library(raster)

# Data
Trees <- genTrees()
Species <- genSpecies()
Env <- genEnv()
com <- quadrats()
com@data$id <- paste0('C', com@data$id)

# L matrix
Trees$SpCode <- tolower(Trees$SpCode)
Trees$ID <- row.names(Trees)
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
XY <- XY[-which(is.na(XY$X)),] # Removing NA
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
Lm[is.na(Lm)] <- 0

# R matrix
com@data$Elevation <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'Elevation'), c('slope', 'curvature', 'cosaspect'))
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
names(com@data)[5] <- 'SW'
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
Rm <- com@data[-1]
row.names(Rm) <- com@data$id

# Q matrix
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
PFT_sp$SLA <- log(PFT_sp$SLA)
PFT_sp$LA <- log(PFT_sp$LA)
Qm <- PFT_sp[c('SLA', 'LDMC', 'WD', 'Thick', 'LA')]
row.names(Qm) <- PFT_sp$Sp_Code

# Matrices adjustments
Qm <- Qm[intersect(row.names(Qm), names(Lm)),]
Qm <- Qm[-which(is.na(Qm$WD)),]
Lm <- Lm[row.names(Rm),]
Lm <- Lm[intersect(row.names(Qm), names(Lm))]
Lm <- Lm / rowSums(Lm)

# Q plot
plot(t(t(Q$li)/Q$eig[1:2]),pch=16, col = 'lightgrey',
     xlab = paste('Axe 1 -', round(Q$eig[1]/sum(Q$eig)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(Q$eig[2]/sum(Q$eig)*100, 2), '%'))
par(new=T)
XLIM <- c(-max(abs(Q$co[,1])),max(abs(Q$co[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Q$co,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
# legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
       # text.col = c("green", "firebrick"))
Hmisc::subplot(function(){
  b <- barplot((Q$eig / sum(Q$eig)), las = 2, ylim = c(0,range(bstick(7))[2]),
          cex.names = 0.7, axes = T, names.arg = NA) ;
  lines(b, bstick(length(Q$eig)), col = 'red') ;
  legend('topright', c('Eigen values', 'broken stick'), fill = c('grey', 'red'), cex = 0.6,
         box.col = 'white')},
  x=grconvertX(c(0.65,0.95), from='npc'),
  y=grconvertY(c(0.01,0.26), from='npc')
)
axis (3)
axis (4)
# arrows (rep(0,nrow(R$co)),rep(0,nrow(R$co)),R$co[,1],R$co[,2],lwd=1, col = 'blue')
# text(R$co, pos = c(4, 2, 4, 2, 3), rownames(R$co),cex=0.9,col='blue')
arrows (rep(0,nrow(Q$co)),rep(0,nrow(Q$co)),Q$co[,1],Q$co[,2],lwd=2)
text (Q$co, pos = c(2, 2, 2, 4, 4), rownames(Q$co),cex=1.2)
```

@@

Multi-colinearity was avoided by reducing our explanatory variables to slope, curvature, wetness and southwesterness, following PCA results (see figure 2). Curvature covary with profile curvature, plan curvature, distance to potential rivers and the PCA first axis. Eleveation (DEM) and wetness are opposed in the principal component analysis and slope and south-westerness doesn't seem to covary much with other variables. Structural equation modelling (SEM) revealed that environmental biotic factors were influencing functionnal traits and not the contrary. Effectively retained model was the first one (**m1** *Traits ~ Abiotic + Biotic*, see supplementary material S1) with a significantly lower AIC on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$)

Whereas total co-inertia from RLQ analysis results was significanlty different for the model *m2*, it didn't differ from null model *m4* in model *m6* proposed by @ter_braak_improved_2012. So because both test should be rejected to reject model *m6* hypothesis, therefore the null hypothesis *H0* is not rejected.

Still fourth-corner analysis with adjusted p-values or not show significant trait-environment association at a level $\alpha = 0.05$ (see table 7). WD is negatively associated with southwesterness for all L matrices and positively associated with canopy height in presence absence. LDMC and LA are respectively negatively and positively associated with canopy height in relative abundance.

Table 7: Summary of fourth-corner analysis. *Result of Monte-Carlo test between fourth-corner result and 999 nul models following the model m6 suggested by @ter_braak_improved_2012.* _L matrices are filled with relative abundance (AB), presence-absence (PA) or relative basal area (BA)). The 5 functionnal traits are tested with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA). See table 1 for trait abbreviations. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

```{r FQR}
library(ade4)
rd <- randtest(rlq, modeltype = 6)

# L matrices list
Lm.l <- list(abundance = Lm)
Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
Lm[is.na(Lm)] <- 0
Lm <- Lm[row.names(Rm),]
Lm <- Lm[intersect(row.names(Qm), names(Lm))]
Lm[Lm > 0] <-  1
Lm <- Lm / rowSums(Lm)
Lm.l$`presence-absence` = Lm
Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum))
Lm[is.na(Lm)] <- 0
Lm <- Lm[row.names(Rm),]
Lm <- Lm[intersect(row.names(Qm), names(Lm))]
Lm <- Lm / rowSums(Lm)
Lm.l$`basal area` = Lm

# Full # -WD
RLQm.l <- list(full = list(Rm.l = rep(list(Rm),3), Lm.l = Lm.l, Qm.l = rep(list(Qm),3)))
Qm <- PFT_sp[c('SLA', 'LDMC', 'Thick', 'LA')]
Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
Lm[is.na(Lm)] <- 0
row.names(Qm) <- PFT_sp$Sp_Code
Qm <- Qm[intersect(row.names(Qm), names(Lm)),]
Lm <- Lm[row.names(Rm),]
Lm <- Lm[intersect(row.names(Qm), names(Lm))]
Lm <- Lm / rowSums(Lm)
Lm.l <- list(abundance = Lm)
Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
Lm[is.na(Lm)] <- 0
Lm <- Lm[row.names(Rm),]
Lm <- Lm[intersect(row.names(Qm), names(Lm))]
Lm[Lm > 0] <-  1
Lm <- Lm / rowSums(Lm)
Lm.l$`presence-absence` = Lm
Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum))
Lm[is.na(Lm)] <- 0
Lm <- Lm[row.names(Rm),]
Lm <- Lm[intersect(row.names(Qm), names(Lm))]
Lm <- Lm / rowSums(Lm)
Lm.l$`basal area` = Lm
RLQm.l$`-WD` <- list(Rm.l = rep(list(Rm),3), Lm.l = Lm.l, Qm.l = rep(list(Qm),3))
# str(RLQm.l,3)

# FQR
fqr <- lapply(RLQm.l, function(x){mapply(function(r,l,q){fourthcorner(r, l, q, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")}, r = x$Rm.l, l = x$Lm.l, q = x$Qm.l, SIMPLIFY = F)})
names(fqr$full) <- names(Lm.l) ; names(fqr$`-WD`) <- names(Lm.l)

# Result
# par(mfrow = c(2,3))
# lapply(fqr, function(x){lapply(x, plot)})
fqr.df.l <- mapply(function(y,z){
  fqr.df <- with(y$tabG, cbind(data.frame(t(matrix(unlist(strsplit(names, ' / ')), nrow = 2)), stringsAsFactors = F), obs, pvalue))
  fqr.wd <- with(z$tabG, cbind(data.frame(t(matrix(unlist(strsplit(names, ' / ')), nrow = 2)), stringsAsFactors = F), obs, pvalue))
  fqr.df <- rbind(fqr.df, fqr.wd[which(fqr.wd$X2 == 'WD'),])
  names(fqr.df)[1:2] <- c('Factor', 'Trait') 
  fqr.df$res <- paste(round(fqr.df$obs, 3), '/', round(fqr.df$pvalue, 3), unlist(lapply(as.list(fqr.df$pvalue), stars, ns = ' ')))
  fqr.df <- tapply(fqr.df$res, INDEX = list(fqr.df$Trait, fqr.df$Factor), paste, simplify = F)
  return(fqr.df)
  }, y = fqr$`-WD`, z = fqr$full, SIMPLIFY = F)
```
```{r FQR abundance}
kable(t(fqr.df.l$abundance))
```

## Are phylogenetic composition of communities linked to the environment ?

Only Thick, LDMC, WES and at a fewer level LES shown a phylogenetic signal positively structured in the phylogenetic tree at a phylogenetic distance between 0 and 100 (see supplementary material S3). $\beta st$ and $\pi st$ indicates both for respectivelly abundance and presence absence data that our phylogeny is overdispersed inside communities: species within communities are less phylogenetically related on average than are specis from distinct communities. But compared to the model null *1s* suggested by @hardy_characterizing_2007 , the phylogenetic overdispersion is almost null except in recent years (from 50 Myrs). Finally, total inertia from RLQ analysis with tip distances of phylogenetic tree as matrix Q didn't differ from a null model were all phylogenetic tree tips were randomized ($p-value = 0.5344$ with 999 repetitions.)

```{r PhyloMaker}
library(phylosignal)
library(phytools)
library(phylobase)
